Skew-symmetric entropy stable discontinuous Galerkin formulations

### 2018-09-08 docs/skewform/main.tex
updating other experiments and docs

@@

\begin{abstract}
Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.  In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which an SBP property does not hold.  These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.  We conclude by discussing the applicability of these new operators for entropy stable schemes on hybrid meshes.% and non-conforming meshes.  
\end{abstract}

@@

High order methods for the simulation of time-dependent compressible flow have the potential to achieve higher levels of accuracy at lower costs compared to current low order schemes \cite{wang2013high}.  In addition to superior accuracy and efficiency on modern computing architectures, the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.  The same properties also make high order methods attractive for the resolution of unsteady phenomena such as vorticular and turbulent flows, which are sensitive to numerical dissipation \cite{visbal1999high, wang2013high}.  However, while high order methods are often provably stable for wave problems, high order schemes for nonlinear conservation laws have consistently been hampered by problems of instability.

When applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.  In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.  The resulting schemes for nonlinear conservation laws walk a fine line between stability, robustness, and accuracy.  Aggressive stabilization or regularization can result in the loss of high order accuracy, while too little can result in instability \cite{wang2013high}.  Moreover, it can be difficult to determine robust expressions for stabilization paramaters, as parameters which work for one simulation can fail when applied to a different physical regime or discretization setting.

These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.  Proofs of continuous entropy inequalities rely on the chain rule.  In contrast, these discrete entropy inequalities account for the loss of the chain rule due to effects such as quadrature errors, which are incurred when applying polynomially exact quadrature rules to nonlinear and rational integrands within the DG formulation.  These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.  These methods have also been extended to staggered grid \cite{parsani2016entropy}, generalized SBP and Gauss nodes \cite{chan2018efficient} and non-conforming meshes \cite{friedrich2017entropy}.  Entropy stable boundary conditions have also been determined for the compressible Euler and Navier-Stokes equations \cite{parsani2015entropy, svard2018entropy}.

\note{Finish: add Section description.  Talk about situations where the SBP or decoupled SBP property doesn't hold: triangles with reduced surface quadrature, GLL quads with GQ face quadratures, and hybrid couplings.}

\section{Entropy stability for systems of nonlinear conservation laws}

@@

\section{Polynomial approximation spaces}

@@

\section{Quadrature-based matrices and decoupled SBP operators}

%\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$ on $\hat{D}$.  These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.  We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific summation-by-parts properties.

@@

Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.  Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q$.  The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
\begin{gather}
\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:QN}\\
\bm{Q}^i_N  = \LRs{
\begin{array}{cc}
\bm{Q}^i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T \bm{W}_f {\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T \bm{W}_f {\rm diag}(\hat{\bm{n}}_i)\\
-\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q & \frac{1}{2} \bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
\end{array}}.  \nonumber
\end{gather}
When combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
\begin{gather}
f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
\label{eq:dnapprox}
\end{gather}
The approximation can also be interpreted as solving a variational problem.  Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.  Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that, for all $v\in V^N\LRp{\hat{D}}$, 
\begin{align*}
\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
\end{align*}

For sufficiently accurate quadrature rules, the matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.  
\begin{lemma}
Assume that the volume and surface quadrature rules are sufficiently accurate such that the quantities
\[
\int_{\hat{D}} \pd{u}{x_i} v, \qquad \int_{\partial \hat{D}} u v n_i
\]
are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ satisfies a summation by parts property:
\begin{gather}
\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\label{eq:dsbp}
\end{gather}
%\begin{enumerate}
%\item The mass matrix is positive definite, 
%\item The quadrature integrates exactly $\int_{\hat{D}} \pd{u}{x_i} v$ for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
%\end{enumerate}
\label{lemma:dsbp}
\end{lemma}
\begin{proof}
The proof is a straightforward extension of Theorem 1 in \cite{chan2017discretely} to polynomial approximation spaces on non-simplicial elements.  
\end{proof}

The assumptions of Lemma~\ref{lemma:dsbp} are satisfied for sufficiently accurate volume and surface quadratures.  For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.  Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if both the volume and surface quadratures are exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ with respect to one coordinate and degree $N$ with respect to others.

@@

\section{Skew-symmetric entropy conservative formulations}

While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.  This difficulty is a result of the choices of volume and surface quadrature which naturally arise in these settings.  We first illustrate how specific pairings of volume and surface quadratures can result in the loss of the SBP property (\ref{eq:dsbp}).  We then propose a skew-symmetric formulation which is entropy conservative without explicitly requiring operators which satisfy the SBP property.

@@

In this section, we give examples of specific pairings of volume and surface quadratures under which the decoupled SBP property does not hold (see Figure~\ref{fig:sbploss}).  We consider two dimensional reference elements $\hat{D}$ with spatial coordinates $x,y$.
\begin{figure}
\centering
\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{figs/gllgauss.png}\label{subfig:gllgq}}
\hspace{2em}
\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Lemma~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}). }
\label{fig:sbploss}
\end{figure}

@@

\begin{figure}
\centering
\begingroup
\captionsetup[subfigure]{width=.425\textwidth}
\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
\endgroup
\hspace{2em}
\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
\caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}).  }
\label{fig:hybrid}
\end{figure}

@@

The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.  We will relax accuracy conditions on these quadrature rules by utilizing not the strong SBP property, but a weaker ``variational'' version of the SBP property.  This variational property relies on the assumption that the volume and surface quadrature rules are exact for polynomials of certain degrees.

\begin{assumption}
\label{ass:quad}
Let $v(\bm{x})$ be a polynomial such that $v \in V^{M_{\rm vol}}$ with surface trace $\LRl{v}_{\partial \hat{D}} \in V^{M_{\rm face}}\LRp{f}$ for $f \in \partial \hat{D}$, where $M_{\rm vol}, M_{\rm face} \leq N$ are two non-negative integers.  We assume that the volume and surface quadrature rules are sufficiently accurate such that
\begin{enumerate}
\item the mass matrix $\bm{M}$ is positive definite, 
\item the volume and surface quadratures exactly integrate the quantities $\int_{\hat{D}} \pd{u}{x_i} v$ and $\int_{\partial \hat{D}} u v n_i$ for all $u \in V^N\LRp{\hat{D}}$, $i = 1,\ldots, d$, and $f \in \partial \hat{D}$.  
\end{enumerate}
\end{assumption}

This section presumes that Assumption~\ref{ass:quad} is satisfied for general  $M_{\rm vol}, M_{\rm surf}$.  In Sections~\ref{sec:affine} and \ref{sec:curved}, specific values of $M_{\rm vol}, M_{\rm surf}$ will be motivated by proofs of entropy stability, and Assumption~\ref{ass:quad} will be verified for specific choices of volume and surface quadrature on simplicial and tensor product elements.
%Suppose Assumption~\ref{ass:quad} is satisfied, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ at volume quadrature points.  Then, for an arbitrary vector $\bm{u}_q$, 
%\[
%\bm{v}_q^T \bm{Q}_i \bm{u}_q = \int_{\hat{D}} \pd{u}{x_i} v, \qquad \LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f \diag{\hat{\bm{n}_i}} \bm{V}_f\bm{P}_q\bm{u}_q =  \int_{\partial \hat{D}} \Pi_N u v n_i,
%\]
%where we have used that, because $v$ is a polynomial of degree $N$, it is equal to its own $L^2$ projection.

>>>>

%\documentclass[preprint,10pt]{elsarticle}
%\usepackage{amsmath,amssymb,amsfonts,amsthm}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}
%\theoremstyle{lemma}
%\newtheorem{lemma}{Lemma}
%\newtheorem*{remark}{Remark}
%\theoremstyle{theorem}
%\newtheorem{theorem}{Theorem}
%\theoremstyle{assumption}
%\newtheorem{assumption}{Assumption}
%\usepackage{fullpage}

@@

\begin{abstract}
Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.  In this work, we show how to construct skew-symmetric schemes which are entropy stable even for volume and surface quadratures under which an SBP property does not hold.  These skew-symmetric formulations avoid the use of a ``strong'' SBP property, and require only that operators exactly differentiate constants and satisfy a discrete form of the fundamental theorem of calculus.   We conclude by discussing mortar-based implementations of skew-symmetric formulations for entropy stable schemes on hybrid meshes. 
\end{abstract}

%\tableofcontents

@@

High order methods for the simulation of time-dependent compressible flow have the potential to achieve higher levels of accuracy at lower costs compared to current low order schemes \cite{wang2013high}.  In addition to superior accuracy, % and efficiency on modern computing architectures, 
the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.  The same properties also make high order methods attractive for unsteady phenomena such as vorticular and turbulent flows, which are sensitive to numerical dissipation \cite{visbal1999high, wang2013high}.

However, 
%while high order methods are often provably stable for wave problems, high order schemes for nonlinear conservation laws have consistently been hampered by problems of instability.  
when applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.  In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.  The resulting schemes for nonlinear conservation laws walk a fine line between stability, robustness, and accuracy.  Aggressive stabilization or regularization can result in the loss of high order accuracy, while too little can result in instability \cite{wang2013high}.  Moreover, it can be difficult to determine robust expressions for stabilization paramaters, as parameters which work for one simulation can fail when applied to a different physical regime or discretization setting.

These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.  Proofs of continuous entropy inequalities rely on the chain rule, which does not hold discretely due to effects such as quadrature error.   By using an approach referred to as ``flux differencing'', entropy stable schemes account for the loss of the chain rule while maintaining a semi-discrete analogue of the continuous entropy inequality.  These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.  Entropy stable methods have also been extended to a variety of other discretization settings, including staggered grids \cite{parsani2016entropy}, Gauss-Legendre collocation \cite{chan2018efficient}, and non-conforming meshes \cite{friedrich2017entropy}.

Entropy stable discretizations are built upon flux differencing and a summation-by-parts (SBP) property, which serves as a discrete analogue of integration by parts for quadrature-based discretization matrices.  However, the SBP property does not hold for some under-integrated quadrature rules, which naturally arise in certain discretization settings.  For example, on hybrid meshes consisting of both quadrilateral and triangular elements, it is convenient to utilize the same quadrature rule on shared faces between different element types.  On degree $N$ tensor product elements, a popular choice of quadrature is an $(N+1)$-point Gauss-Legendre-Lobatto (GLL) rule.  When both volume and surface integrals are approximated using $(N+1)$ point GLL quadrature rules, the SBP property holds, despite the fact that GLL quadrature is inexact for the integrands which appear in finite element formulations \cite{fisher2013high}.  However, while GLL quadrature induces an SBP property on quadrilateral elements, but does not always result in an SBP property if used for triangular elements \cite{chan2017discretely}.  %A similar issue arises on non-conforming meshes, where a composite ``split-side'' quadrature naturally when coupling larger elements to smaller elements through a non-conforming interface \cite{kozdon2018energy}.  For quadrilaterals utilizing a GLL volume quadrature, the use of a composite quadrature on a non-conforming face also results in the loss of the SBP property \cite{chan2017discretely}.

This work proposes an alternative formulation which utilizes a skew-symmetric form of flux differencing.  Under such a formulation, the proof of entropy stability requires only a weaker ``variational'' form of the SBP property, which holds under a more general class of quadrature rules compared to the decoupled SBP property introduced in \cite{chan2017discretely, chan2018discretely}.  We show that the skew-symmetric formulation is entropy stable, locally conservative, and free-stream preserving on curved elements, and confirm theoretical results with numerical experiments on triangular, quadrilateral, and hybrid triangular-quadrilateral meshes.

It should be noted that a similar approach to entropy stable discretizations was introduced within a finite difference framework \cite{chen2017entropy, crean2018entropy} using multidimensional differencing operators which are constructed to satisfy accuracy conditions and an SBP property \cite{hicken2016multidimensional}.  These operators do not correspond to any specific basis or approximation space, but can be shown to exist for nodal points corresponding to sufficiently accurate choices of volume and surface quadrature.  The formulations presented in this work require similar accuracy conditions, but differ from those of \cite{chen2017entropy, crean2018entropy} in that they build upon an explicit approximation space.  In contrast with traditional SBP coupling terms \cite{crean2017high}, the use of decoupled SBP operators also accomodates general choices of volume and surface quadrature (e.g.\ volume quadratures without boundary nodes and over-integrated rules) while ensuring compact coupling conditions between neighboring elements.

The structure of the paper is as follows: Section~\ref{sec:nonlin} describes the continuous entropy inequality which we aim to replicate discretely.  Section~\ref{sec:approx} and Section~\ref{sec:sbp} introduce polynomial approximation spaces and quadrature-based SBP operators on simplicial and tensor product elements.  Section~\ref{sec:skew1} then introduces the variational SBP property and an entropy conservative skew-symmetric formulation.  Section~\ref{sec:skew2} extends the skew-symmetric formulation to curved elements, and Section~\ref{sec:num} presents numerical experiments which verify the theoretical assumptions, stability, and accuracy of the proposed formulations.

%\note{Finish: add Section description.  Talk about situations where the SBP or decoupled SBP property doesn't hold: triangles with reduced surface quadrature, GLL quads with GQ face quadratures, and hybrid couplings.}

\section{Entropy stability for systems of nonlinear conservation laws}
\label{sec:nonlin}

@@

\section{Polynomial approximation spaces}
\label{sec:approx}

@@

\section{Quadrature-based matrices and decoupled SBP operators}
\label{sec:sbp}
%\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$ on $\hat{D}$.  These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.  We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific summation-by-parts properties.

@@

Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.  Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q$.  The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
\begin{gather}
\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}}, \label{eq:QN}\\
\bm{E} = \bm{V}_f\bm{P}_q, \qquad \bm{B}^i = \bm{W}_f \diag{\hat{\bm{n}}_i}, \qquad \bm{Q}^i_N  = \LRs{
\begin{array}{cc}
\bm{Q}^i - \frac{1}{2}\bm{E}^T\bm{B}^i\bm{E} &  \frac{1}{2}\bm{E}^T\bm{B}^i\\
-\frac{1}{2}\bm{B}^i\bm{E} & \frac{1}{2} \bm{B}^i
\end{array}}.  \nonumber
\end{gather}
Here, $\bm{B}^i$ is a boundary matrix, and $\bm{E}$ denotes the extrapolation matrix which maps values at volume quadrature points to values at surface quadrature points using quadrature-based $L^2$ projection and polynomial interpolation.

It was shown in \cite{chan2017discretely, chan2018efficient} that, when combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
\begin{gather}
f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
\label{eq:dnapprox}
\end{gather}
%The approximation can also be interpreted as solving a variational problem.  Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.  Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that, for all $v\in V^N\LRp{\hat{D}}$, 
%\begin{align*}
%\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
%\end{align*}

For sufficiently accurate quadrature rules, the matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.  
\begin{lemma}
Assume that the volume and surface quadrature rules are sufficiently accurate such that the quantities
\[
\int_{\hat{D}} \pd{u}{x_i} v, \qquad \int_{\partial \hat{D}} u v n_i
\]
are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ (\ref{eq:QN}) satisfies a summation by parts property:
\begin{gather}
\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \bm{B}^i_N, \qquad \bm{B}^i_N = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{B}^i\end{array}}.\label{eq:dsbp}
\end{gather}
%\begin{enumerate}
%\item The mass matrix is positive definite, 
%\item The quadrature integrates exactly $\int_{\hat{D}} \pd{u}{x_i} v$ for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
%\end{enumerate}
\label{lemma:dsbp}
\end{lemma}
\begin{proof}
The proof is a straightforward extension of Theorem 1 in \cite{chan2017discretely} to polynomial approximation spaces on non-simplicial elements.  
\end{proof}

The assumptions of Lemma~\ref{lemma:dsbp} are satisfied for sufficiently accurate volume and surface quadratures.  For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.  Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if both the volume and surface quadratures are exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ polynomials with respect to one coordinate and degree $N$ with respect to others.

@@

\section{Skew-symmetric entropy conservative formulations on a single element}
\label{sec:skew1}
While the SBP property has been used to derive entropy stable schemes, the strong SBP property (\ref{eq:dsbp}) is difficult to enforce in certain discretization settings, such as hybrid and non-conforming meshes.  This difficulty is a result of the choices of volume and surface quadrature which naturally arise in these settings.  We first illustrate how specific pairings of volume and surface quadratures can result in the loss of the SBP property (\ref{eq:dsbp}).  We then propose a skew-symmetric formulation which is entropy conservative without explicitly requiring operators which satisfy the SBP property.

@@

In this section, we give examples of specific pairings of volume and surface quadratures under which the decoupled SBP property does not hold (see Figure~\ref{fig:sbploss}).  We consider two dimensional reference elements $\hat{D}$ with spatial coordinates $x,y$.
\begin{figure}
\centering
\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{figs/gllgauss.png}\label{subfig:gllgq}}
\hspace{2em}
\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Lemma~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}). Volume quadrature nodes are drawn as circles, while surface quadrature nodes are drawn as squares.}
\label{fig:sbploss}
\end{figure}

@@

\begin{figure}
\centering
\begingroup
\captionsetup[subfigure]{width=.425\textwidth}
\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
\endgroup
\hspace{2em}
\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
\caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}). Volume quadrature nodes are drawn as circles, while surface quadrature nodes are drawn as  squares. }
\label{fig:hybrid}
\end{figure}

@@

The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.  We will relax accuracy conditions on these quadrature rules by utilizing a weaker ``variational'' version of the SBP property variational, which relies on the exactness of volume and surface quadrature rules only for specific polynomials.

On tensor product elements, we restrict ourselves to isotropic volume quadrature rules which are construced from tensor products of one-dimensional quadrature formulas.  For the remainder of this work, the degree of the multi-dimensional quadrature rule on tensor product elements will refer to the degree of exactness of the one-dimensional rule.  For example, we refer to the quadrature rule constructed through a tensor product of one-dimensional $(N+1)$-point GLL quadrature rules as a degree $(2N-1)$ quadrature rule.  This choice of quadrature is sufficient to guarantee that the mass matrix is positive definite \cite{canuto2007spectral}.

Throughout the remainder of this work, we make the following assumptions:
\begin{assumption}
\label{ass:quad}
Let $v \in V^{N}$ denote a fixed polynomial.  We assume that: 
\begin{enumerate}
\item the mass matrix $\bm{M}$ is positive definite under the volume quadrature rule,
\item the volume quadrature rule is exact for integrals of the form\\$\int_{\hat{D}} \pd{u}{\hat{x}_j} v$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$.
\item the surface quadrature rule is exact for integrals of the form\\$\int_{\partial \hat{D}} u v n_j$ for all $u \in V^N\LRp{\hat{D}}$, $j = 1,\ldots, d$, and $f \in \partial \hat{D}$.  
\end{enumerate}
\end{assumption}

The conditions of Assumption~\ref{ass:quad} are relatively standard within the SBP literature \cite{hicken2016multidimensional, chan2017discretely, crean2018entropy}, though they have not previously depended on the specific choice of polynomial $v(\bm{x})$.   The strong SBP property (\ref{eq:dsbp}) can be derived if Assumption~\ref{ass:quad} is satisfied for all $v \in V^N$.  In contrast, we will derive a weaker SBP property which assumes only that Assumption~\ref{ass:quad} holds for a few specific polynomials $v$.  In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific polynomials $v$ will be motivated by proofs of entropy stability, and we will present examples of volume and surface quadrature rules on simplicial and tensor product elements which satisfy Assumption~\ref{ass:quad} for these choices of $v$.  
% the quantities $\int_{\hat{D}} \pd{u}{x_i} v$ and

%This section presumes that Assumption~\ref{ass:quad} is satisfied for general  $M_{\rm vol}, M_{\rm surf}$.  In Sections~\ref{sec:singleelem} and \ref{sec:curved}, specific values of $M_{\rm vol}, M_{\rm surf}$ will be motivated by proofs of entropy stability, and Assumption~\ref{ass:quad} will be verified for specific choices of volume and surface quadrature on simplicial and tensor product elements.

The conditions of Assumption~\ref{ass:quad} imply the following variational SBP property:
\begin{lemma}
\label{lemma:vsbp}
%Assume that $v \in V^M$.  
%If the volume quadrature is exact for polynomials of degree $N+M-1$ and the surface quadrature is exact for polynomials of degree $N+M$, 
Suppose Assumption~\ref{ass:quad} holds, and let $\bm{v}_q$ denote the vector of values of $v(\bm{x})$ of at volume quadrature points. Let $u(\bm{x})$ be an $L^2$ integrable function, and let $\bm{u}_q$ denote the values of $u(\bm{x})$ at quadrature points.  Then, 
\[
\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ {\bm{E}}^T \bm{B}^i\bm{E} - \bm{Q}_i^T}\bm{u}_q.
\]
\end{lemma}
\begin{proof}
Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.  The quadrature-based $L^2$ projection $\Pi_Nu$ is a polynomial of degree $N$, and is computed discretely as $\bm{P}_q\bm{u}_q$.  By the exactness of the volume quadrature in Assumption~\ref{ass:quad}, 
\begin{align*}
\bm{v}_q^T\bm{Q}_i \bm{u}_q &= \bm{v}_q^T\bm{W} \bm{D}^i \bm{P}_q \bm{u}_q = \int_{\hat{D}} \pd{\Pi_N u}{\hat{x}_i} v = \int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i - \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i}.
\end{align*}
%Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{N+M-1}$, it is exactly integrated using the volume quadrature.  
%Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and 
By the exactness of the surface quadrature in Assumption~\ref{ass:quad}, 
\begin{align*}
&\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} =\\
 &\bm{v}_f^T\bm{B}^i \bm{E}\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
=& \bm{v}_q^T {\bm{E}}^T\bm{B}^i \bm{E}\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
\end{align*}
where we have used that, since $v \in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q = \bm{E}\bm{v}_q$.  The proof is completed by noting that 
\[
\bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{M}\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T\bm{W} = \bm{Q}_i^T.
\]
\end{proof}
